---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# HFSutils

Personal helper functions for statistical modeling, simulation, and reporting in applied research.

This package collects small, reusable functions that I use across projects to avoid
code duplication and to keep analysis scripts readable. Functions are documented but
primarily designed for personal use rather than as a general-purpose public API.

## Scope and philosophy

- Functions are **opinionated** and reflect my own workflows.
- APIs may change over time.
- Not all edge cases are handled.
- For published work, package versions should be **pinned** for reproducibility.

The package is organized by functionality rather than by project.


## Installation

You can install the development version of HFSutils from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("hafsunde/HFSutils")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(HFSutils)

SEtoCI(
  estimates = c(0.2, 0.4),
  std_errors = c(0.05, 0.05)
)


```


## Contents

The package currently includes helpers for:

### Simulation
- Exact control of empirical variance
- Adding noise to reach a target variance

### Formatting
- Numeric formatting for tables and figures
- Percent, integer, and scientific notation helpers

### Scale scoring
- Reverse coding of questionnaire items
- Missing-data tolerant score computation

### OpenMx utilities
- Extraction of estimates, standard errors, and Wald confidence intervals

Some helper functions are intentionally internal and accessed via `HFSutils:::`.

## Function overview (for newcomers)

If you're new to this package, this list shows what each exported function is for and when to use it.

### Confidence intervals and correlations

- `SEtoCI()` — Convert estimates + standard errors into confidence intervals; supports Fisher z back-transformation for correlations.
- `cor_fastCI()` — Compute a correlation and a quick Fisher z confidence interval from two vectors.

### Simulation helpers

- `exact_rnorm()` — Simulate normal data and (optionally) rescale so the sample SD exactly matches a target SD.
- `noise_filler()` — Add random noise to a variable when its variance is below 1, aiming for unit variance.

### Education recoding helpers

- `convert_edu_level_to_years()` — Recode NUS2000 education levels (0–8) to approximate years of education.
- `convert_edu_level_to_group()` — Recode NUS2000 education levels into broad education groups.

### Score construction

- `score_scale()` — Build scale scores from item-level data with optional reverse coding and missingness thresholds.

### Plot helper

- `HFS_theme()` — Apply a `ggplot2` theme with transparent background and customizable foreground color.

### Number formatting helpers

- `numformat()` — General numeric formatter with commas, decimal control, and optional leading-zero removal.
- `format_num()` — Force fixed decimal formatting (e.g., `1.20`).
- `format_int()` — Format integer-like values with thousands separators.
- `to_percent()` — Convert proportions to percent strings.
- `expSup()` — Format scientific notation as LaTeX-style exponent text.
